---
title: "Config files"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Config files}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#"
)
```

## Purpose

This vignette provides an overview of how to use a config file with the {aftables} package to set workbook properties and formatting.

## Workbook properties and formatting which can be set in a config file

Workbook properties:

 - Author
 - Title
 - Keywords
 - Subject
 - Category
 - Comments

Analysis Function guidance advises workbooks should have the author, title, keywords and language document properties completed. aftables provides functionality to set the author, title and keywords properties. See ['Releasing statistics in spreadsheets'](https://analysisfunction.civilservice.gov.uk/policy-store/releasing-statistics-in-spreadsheets/) for more information including how to set the workbook language.

Workbook formatting:

 - Font
 - Base font size
 - Table header font size
 - Sheet heading font size
 - Sheet subheading font size
 - Default column widths
 - Wider column widths
 - Text wrapping

## Structure of a config file

The config file must include an `aftables` entry. {aftables} will only look for workbook configs below the `aftables` entry. The config file may include any other entries. This allows for {aftables} config files to be combined with other config files for multiple purposes.

The `aftables` config file supports multiple entries. There must be at least one config below the `aftables` entry, which could be named `default` or be user-specified. 

Each config must contain at least one of `workbook_properties` and `workbook_format`. Each value below `workbook_properties` must be a character string apart from `keywords`, which is a list. The `base_font_name` value of `workbook_format` must be a character string, and each other value must be an integer. {aftables} checks all values to ensure they are character strings or integers as required to process them correctly.

This is the contents of the example config file included with {aftables}:

```{r results='asis', echo = FALSE}
cat('```xml\n')
cat(readLines(system.file("ext-data",
                          "config.yaml",
                          package = "aftables")),
    sep = "\n")
cat('\n```\n')
```

The `create_config_yaml()` function outputs a copy of this config file to a location of the users choice. The default location is the current working directory. We recommend using the example config file as a basis for developing your own {aftables} config file.

## Applying a config file with `generate_workbook()`

The `generate_workbook()` function has two arguments for config files: `config_path` and `config_name`. Both arguments are optional, and {aftables} will take different action based on which arguments are provided.

If neither `config_path` or `config_name` is provided (the default behaviour), {aftables} will look in the current working directory for a config file named `config.yaml`, and will look in that config file for an `aftables` entry with a `default` config to process.

If `config_path` is provided without `config_name`, {aftables} will use `config_path` to find a config file, and will look in that config file for an `aftables` entry to process. The `default` config will be processed and its `workbook_properties` and `worbook_format` values applied to the workbook created.

If `config_name` is provided without `config_path` {aftables} will look in the current working directory for a config file named `config.yaml`, and will look in that config file for an `aftables` entry to process. {aftables} will look for a custom config matching `config_name`, as well as any `default` config. A combined config is generated from the `default` config and the `config_name` custom config, with values specified in the `config_name` custom config preferred over the `default` config. If no `default` config exists all values are taken from the `config_name` custom config.

If both `config_path` and `config_name` are provided, {aftables} will use `config_path` to find a config file, and will look in that config file for an `aftables` entry to process. {aftables} will look for a custom config matching `config_name`, as well as any `default` config. A combined config is generated from the `default` config and the `config_name` custom config, with values specified in the `config_name` custom config preferred over the `default` config. If no `default` config exists all values are taken from the `config_name` custom config.

Example using the `workbook1` config from the {aftables} example config file:

```{r create-af-wb, eval = FALSE}
my_wb <- aftables::generate_workbook(demo_aftable,
                                     config_path = system.file("ext-data",
                                                               "config.yaml",
                                                               package = "aftables"),
                                     config_name = "workbook1")
```

Running this code would produce an {openxlsx2} wbWorkbook object named `my_wb` using the `workbook1` config in the {aftables} example config file. The `default` config and the `workbook1` config would be combined. The `category` value in the `default` config would not be used, and instead the `category` value from the `workbook1` config would be used. The `keywords` value from the `workbook1` config is appended to the 3 values from the `default` config.
